\documentclass[11pt]{book}
\usepackage{res_doc_log}
\usepackage[ruled]{algorithm2e}
\usepackage{algorithmic}
\usepackage{amsmath} 
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{booktabs} 
\usepackage[numbers,sort]{natbib}
\usepackage{url} 
\usepackage{textcomp}
\usepackage{mathtools}
\usepackage{caption}
\usepackage{longtable}

\thesistitle{A combined resource allocation and epidemic model to evaluate UAV vaccine delivery and team allocation strategies for measles outbreak response}
\student{Dean Michael Matter}
\graduation{December}
\myyear{2019}
\frontmatter

%Study leader
\supervisor{Dr Linke Potgieter}

\begin{document}
\frontpage[honsbcom]{or}
\dominitoc 
\engabstract{
Humanitarian aid in developing countries is often faced with a wide array of challenges, one of which being the lack of sufficient and stable road infrastructure. In the context of an epidemic, this may result in inefficient outbreak responses such as the late delivery of vaccines. Unmanned aerial vehicles have emerged as a potential alternative to land-based delivery methods, and have been successfully used for an ever-increasing number of medical deliveries in recent years. In this project, the effectiveness of unmanned aerial vehicles for vaccine delivery (in comparison to land-based delivery) in the context of rural epidemic outbreaks (with a particular focus on measles epidemics), is evaluated, along with various vaccine and team allocation strategies.\\
A combined resource allocation and epidemic model is developed to simulate vaccine delivery during a measles outbreak. A network approach is followed in this simulation model, with the network consisting of local populations connected by the migration of individuals. The measles epidemic is simulated by using a compartmental model, which comprises of a system of difference equations for each local population, with a small proportion of each population migrating to other local populations daily. During each time step, optimised daily vaccine delivery schedules and team allocations are determined, by modelling each as a knapsack problem. The objective of each is to minimise the expected number of people exposed to measles, and consequently, minimise the number of measles-induced deaths. Heuristic approaches for resource allocation are also considered. It is clear from simulations that faster outbreak response has a large impact on the number of deaths prevented. Furthermore, finding optimised daily team allocations resulted in a 22.4\% reduction in simulated vaccine and delivery costs. Unmanned aerial vehicles are shown to be capable of meeting demand for vaccines in a simulated measles outbreak, and reduce costs and deaths when used instead of a land-based delivery network in areas with poor road infrastructure.
}

\mainmatter
\tableofcontents
%\listoffigures
%\listoftables

\listofsymbols
\begin{longtable}[c]{ll}
Symbol & Description \\ \hline
\endfirsthead
%
\endhead
%
$n$ & Number of locations in the network. \\
$\mathcal{T}$ & Number of days simulated. \\
$\mathcal{H}$ & Number of working minutes per day. \\
$\mathcal{N}$ & Total number of vaccination teams. \\
$\eta_{i,t}$ & Number of teams at location $i$ on day $t$. \\
$\psi_{i}$ & Number of vaccinations that each team in location $i$ can perform daily. \\
$\tau_{i,t}$ & Number of people present to be vaccinated, on day $t$ at location $i$. \\
$\nu_{i,t}$ & Number of vaccines on-hand on day $t$ at location $i$. \\
$v_{i,t}$ & Number of vaccinations performed on day $t$ at location $i$. \\
$\theta^{S}_{i,t}$ & Proportion of remaining unvaccinated population that is susceptible, \\ &on day $t$ at location $i$. \\
$\theta^{E_{72}}_{i,t}$ & Proportion of remaining unvaccinated population that has been exposed \\ &in the preceding 72 hours, on day $t$ at location $i$. \\
$\lambda_{S}$ & Vaccine success rate in susceptible individuals. \\
$\lambda_{E}$ & Vaccine success rate in individuals exposed for less than 72 hours. \\
$S_{i,t}$ & Number of susceptible individuals at location $i$ on day $t$. \\
$E_{i,t}$ & Number of exposed individuals at location $i$ on day $t$. \\
$I_{i,t}$ & Number of infectious individuals at location $i$ on day $t$. \\
$R_{i,t}$ & Number of recovered individuals at location $i$ on day $t$. \\
$V_{i,t}$ & Number of vaccinated individuals at location $i$ on day $t$. \\
$D_{i,t}$ & Number of deceased individuals at location $i$ on day $t$. \\
$\mathcal{D}_{i,j}$ & Distance by direct road between locations $i$ and $j$. \\
$\Delta_{i,j}$ & Euclidean distance between locations $i$ and $j$. \\
%$i^{*}$ & Location in the network at which the Distribution Centre is placed. \\
$T_{i,j}$ & Number of individuals moving from location $i$ to $j$ daily. \\
$m_{i,j}$ & Proportion of population at location $i$ moving to location $j$ daily. \\
$v'_{i,t}$ & Expected number of vaccines still required at location $i$ on day $t$. \\
$\delta_{d,i}$ & Total flight time for a UAV delivery from the Distribution Centre to location $i$. \\
$\delta_{v,i,j}$ & Total travel time between locations $i$ and $j$, by road. \\
$\mathcal{R}$ & Set of all vehicle delivery routes considered. \\
$\delta_{v,r_{k}}$ & Total travel time for a route $r_{k}$, by road. \\
$E^{*}_{x_{i},i,t+1}$ & Expected prevented exposures resulting from a delivery of $x_{i}$ vaccines to \\ &location $i$ on day $t$. \\
$E^{*}_{r_{k},t+1}$ & Expected prevented exposures resulting from deliveries of $v'_{i,t}$ vaccines to \\ &each location $i$ in route $r_{k}$, on day $t$. \\
$E^{*}_{\eta_{i,t},i,t+1}$ & Expected prevented exposures resulting from $n_{i,t}$ teams being assigned to \\ &location $i$, on day $t$.
\label{tab:symbols_list}\\
\end{longtable}


%\input{sub1_doc_example.tex}

\chapter{Introduction}
\input{intro.tex}

%\chapter{Literature Summaries}
%\input{literaturesummaries.tex}

\chapter{Literature Review}
\input{lit_review.tex}

\chapter{Methodology}
\input{methodology.tex}

\chapter{Results}
\input{results.tex}

\chapter{Conclusion}
\input{conclusion.tex}

\bibliographystyle{orion_and}
%\cleardoublepage
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{refs}
%You need to look at the Referencing.pdf document on Sunlearn to ensure all references are correctly formatted

%\cleardoublepage
\addcontentsline{toc}{chapter}{Appendix}
\appendix
\input{appendix.tex}

\end{document}